---
title: "Homework 2- Statistical Thinking"
author: "Arjun Kumar"
date: "2024-01-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(ggplot2)
library(tibble)
library(rvest)
library(dplyr)
library(readr)
library(stringr)
library(stringi)
```
## **Question 1** ##
```{r echo = FALSE}
prof_data = read.csv("profs.csv")
```
**Part A**

```{r echo = FALSE, message=FALSE}
eval_hist = ggplot(prof_data, aes(x = eval)) + geom_histogram(binwidth = 0.25, fill = "aquamarine", color = "black") + labs(title = "Distribution of Course Evaluation Scores",x = "Evaluation Score",y = "Frequency") + theme_classic()
eval_hist
eval_median = median(prof_data$eval)
eval_mean = mean(prof_data$eval)
```

*Median: `r eval_median` | Mean: `r round(eval_mean,3)`*

This histogram depicts a heavy left skew, demonstrating how most of the evaluation ratings are more lenient towards the 4-5 region in comparison to the low ratings of around 2-3. This is apparent through the median rating, as well as the mode being concentrated at 4, while the mean is 3.998, being very close to the median value as well. This demonstrates how the school has an overall high evaluation rating of the professors, correlating to a high degree of satisfaction with the education the students are receiving.

**Part B**

```{r echo = FALSE, message=FALSE}
prof_native_plots = ggplot(prof_data, aes(x = native, y = eval, fill = native)) + geom_boxplot() + labs(title = "Distribution of Course Evaluation Scores by Native English Speaker",x = "Native English Speaker",y = "Evaluation Score") + scale_fill_manual(values = c("red", "green")) + theme_classic()
prof_native_plots
native_mean = aggregate(eval ~ native, data = prof_data, FUN = mean)
mean_by_native <- aggregate(eval ~ native, data = prof_data, FUN = mean)
mean_table <- as.data.frame(mean_by_native)
kable(mean_table, caption = "Mean Evaluation Scores by Native English Speaker", align = "c")
```

Although there is not a significant difference, it is apparent that the professors who's native language is English have higher evaluation ratings overall compared to those who's native language is not English. This is seen through the overall distribution of the non-native speaker's having a lower Q1, median and Q3 as well as maximum value being lower as well. However, the native speakers do have multiple outliers, including the lowest rating value of the dataset. However, the mean evaluation of the native speakers comes out to 4.02, while the non-native speakers was 3.69, demonstrating a sizeable difference between the two.

**Part C**

```{r echo = FALSE, message=FALSE}
ggplot(prof_data, aes(x = eval)) + geom_histogram(binwidth = 0.39, col = 'black', fill = 'blue') + xlab('Evaluation Score') + ggtitle('Course Evaluation Scores by Gender Distribution') + facet_wrap(~gender) + theme_classic()
mean_by_gender = aggregate(eval~gender, data = prof_data, FUN = mean)
mean_table2 = as.data.frame(mean_by_gender)
kable(mean_table2, caption = "Mean Evaluation Scores by Gender", align = "c")
```
The distribution for female professors' evaluation seems to be more uniform in comparison to the slightly left skewed male distribution, suggesting that men have a slightly higher overall evaluation rating in comparison to women. In order to delve deeper into this case, I created a table showing the means for each gender, and found that the mean of women professors' evaluation was around 3.90, in comparison to the 4.01 men have.

**Part D**

```{r echo = FALSE, message=FALSE}
beauty_plot <- ggplot(prof_data, aes(x = beauty, y = eval)) + geom_point(color = "black", alpha = 0.5) + geom_smooth(method = "lm", se = FALSE, color = "red") +  labs(title = "Scatterplot of Physical Attractiveness vs. Course Evaluations", x = "Physical Attractiveness", y = "Evaluation Score") + theme_classic()
beauty_plot
correlation_coefficient <- cor(prof_data$beauty, prof_data$eval)
```

*Correlation Coefficient: `r correlation_coefficient`*

Looking at the scatterplot and the line of best fit, there is a very slight, positive correlation, nearing no correlation whatsoever. This leads me to believe the attractiveness of the professor has no significant impact on the evaluation they receive. The correlation coefficient is also around 0.18, further proving this point.

## **Question 2** ##
```{r echo=FALSE}
bikes = read.csv("bikeshare.csv")
```

**Data Wrangling**
```{r echo=FALSE, message=FALSE}
rentals_by_hour <- bikes %>%
  group_by(hr) %>%
  summarise(avg_rentals = mean(total))

rentals_by_hours_by_workingday <- bikes %>%
  group_by(hr, workingday) %>%
  summarise(avg_rentals = mean(total))

ridership <- bikes %>%
  filter(hr == 9) %>%
  group_by(weathersit, workingday) %>%
  summarise(avg_rentals = mean(total))
```

**Part A**

```{r echo = FALSE, message=FALSE}
hourly_bike_rentals <- ggplot(rentals_by_hour, aes(x = hr, y = avg_rentals)) + geom_line(color = "blue") + labs(title = "Average Hourly Bike Rentals Throughout Day", x = "Hour", y = "Bike Rental Average") + theme_classic()

hourly_bike_rentals
```

The hourly activity regarding bike rentals seems to spike up exponentially from hour 5 to 8, then drops back down until hour 10. After this, it steadily in creases till around hour 12 then plateaus. Then it sees a slight dip and increase between hours 13 and 14, before spiking up again from hours 15 to 18. It then declines heavily back down until hour 24. The most activity is seen around the middle hours of the day, which makes sense as it aligns with people's schedules the most. This line graph displays the average hourly bike rentals across all hours of the day in Washington D.C. The x-axis represents the hour of the day, while the y-axis represents the average number of bike rentals. The blue line indicates the trend of bike rentals throughout the day.

**Part B**

```{r echo =  FALSE, message=FALSE}
hourly_workingday <- ggplot(rentals_by_hours_by_workingday, aes(x = hr, y = avg_rentals, group = workingday)) + geom_line(color = "blue") + labs(title = "Average Rentals for Bikes by Hour, Faceted by Working Day", x = "Hour", y = "Bike Rental Average") + facet_wrap(~ workingday) + theme_classic()
hourly_workingday
```

On a non-working day(0), the hourly rental average sees a basic trend, going up from hours 5 to approximately 12.5 hours. Then it steadily goes down until hour 24. For a working day(1), The rentals steeply incline until around hour 8, the steeply declines until 10. It then goes steadily up slowly until hour 15, then exponentially upwards till hour 18, then steeply drops off until 24. The schedule for a workday is much more erratic on a workday, potentially due to the work schedule of many people, and the more relaxed schedule of a non-work day. This faceted line graph illustrates the average bike rentals by hour of the day, faceted by whether it is a working day or not. The x-axis shows the hour of the day, and the y-axis represents the average number of bike rentals. The blue lines depict the trend of bike rentals for each category of working day.

**Part C**

```{r echo = FALSE,message=FALSE}
ridership9am <- ggplot(ridership, aes(x = factor(weathersit), y = avg_rentals, fill = factor(weathersit))) + geom_bar(stat = "identity") + labs(title = "Ridership Average at 9am Faceted by Weather Situation", x = "Weather Situation Code", y = "Average Ridership") + facet_wrap(~ workingday) + theme_classic()
ridership9am
```

The weather situations, in which 
1 represents: Clear, Few clouds, Partly cloudy, Partly cloudy
2 represents: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
3 represents:Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog (Not in Graphs)
On weekends, the more favorable the weather conditions, the more ridership was taken seen in the graph with 0. However, on workdays, weather situations 1 and 2 saw almost the same ridership at 9am, but declines off at weather condition 3. This could be due to the fact that these forms of transportation can be necessary for some people to get to work as compared on the weekend.The average ridership during the 9 AM hour is shown in this faceted bar plot by weather scenario code, which is further broken down by whether or not it is a working day. The average ridership is represented by the y-axis, while the weather status code is displayed on the x-axis. The weather status code is reflected in the color of the bars. A working day (1) or a non-working day (0) is represented by each facet. This map sheds light on how working days and the weather impact ridership patterns during morning rush hour.

## **Question 3** ##
```{r echo=FALSE}
capmetro<- read.csv("capmetro_UT.csv")
```

**Re-order variables and Data Wrangling**

```{r echo = FALSE, message=FALSE}
# Reorder categorical variables
capmetro <- mutate(capmetro,
                       day_of_week = factor(day_of_week, levels=c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
                       month = factor(month, levels=c("Sep", "Oct", "Nov")))
boarding_avg <- capmetro %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarise(avg_boardings = mean(boarding))

```

**Part A**

```{r echo = FALSE,message=FALSE}
boarding_average <- ggplot(boarding_avg, aes(x = hour_of_day, y = avg_boardings, color = month)) + geom_line() + facet_wrap(~ day_of_week) +
labs(title = "Boarding Hours Average by Day, Dy of Week, and Month", x = "Hour", y = "Boarding Average", color = "Month") + theme_classic()
boarding_average
```
Regardless of the month, Saturday and Sunday are consistently the least active regarding boardings. All of the weekdays have a very similar trend in which they increase highly, but steadily until evening hours(around 15) then see a steep decline. One thing that I also picked up on is that October consistently seemed to be one of the busiest months regardless of the day of the week. This could be due to factors such as work being the busiest during this month, or other lurking variables. The average number of boardings for the UT bus network in Austin is shown in this faceted line graph by hour of the day, day of the week, and month. Every facet symbolizes a weekday, while the lines signify distinct months (September, October, November). The hour of the day is shown by the x-axis, while the average number of boardings is represented by the y-axis. The graphic sheds light on how ridership patterns vary throughout the course of the week and the months.

**Part B**

```{r echo = FALSE, message=FALSE}
temp_boarding <- ggplot(capmetro, aes(x = temperature, y = boarding, color = weekend)) + geom_point() + facet_wrap(~ hour_of_day) + labs(title = "Boardings vs. Temperature by Hour of the Day", x = "Temperature (F)", y = "Boardings") + theme_classic()
temp_boarding
```

The association between boardings and temperature for the UT bus network in Austin is displayed in this faceted scatter plot each hour of the day. Depending on whether it is a weekday or a weekend, the data points are colored differently. An hour of the day is represented by each facet. The temperature in degrees Fahrenheit is represented by the x-axis, while the number of boardings is represented by the y-axis. We can study how ridership patterns are impacted by temperature throughout the day, taking into account travel on both weekdays and weekends, thanks to this figure. Although the distribution of temperatures is nearly the same whether it is a weekend or weekday, the amount of boardings during the weekdays is consistently much higher than the amount on the weekends.

## **Question 4** ##

```{r echo = FALSE,message=FALSE}
billboard <- read.csv("billboard.csv")
```

**Part A**
```{r echo = FALSE,message=FALSE}
top_10 <- billboard %>%
  count(performer, song, name = "count") %>%
  arrange(desc(count)) %>%
  head(10)
 
top_table <- kable(top_10, format = "markdown", align = c("l", "l", "c"), caption = "10 Most Popular Songs since 1958 Based on Weeks Spent on The Billboard 100") %>%
kable_styling(full_width = FALSE)

top_table
```

This table displays the top 10 songs based on the number of weeks each song appeared on the Billboard Top 100 chart.

**Part B**

```{r echo = FALSE, message=FALSE}
diversity <- billboard %>%
  filter(year != 1958 & year != 2021) %>%
  group_by(year, song) %>%
  summarise() %>%
  group_by(year) %>%
  summarise(unique_songs = n())

ggplot(diversity, aes(x = year, y = unique_songs)) + geom_line() + labs(title = "Diversity of Music in Billboard Top 100 from 1958-2021", x = "Year", y = "Unique Songs") + theme_classic()
```

The musical variety of the Billboard Top 100 chart over time is depicted in this line graph. The number of unique songs that appear at each position on the Billboard Top 100 list in any week for each year is displayed on the y-axis, while the x-axis indicates the years. Most notably, the graph shows patterns in the range of songs included on the list, offering insights into how the music industry has changed over time. The amount of unique songs was very high during the last 1960s to early 1970s, then drastically declines to its lowest point in 2000. It then began to pick back up to near its peak of 800 in 2020.

**Part c**

```{r echo = FALSE,message=FALSE}
ten_week <- billboard %>%
  group_by(performer, song) %>%
  summarise(weeks_on_chart = n_distinct(week)) %>%
  filter(weeks_on_chart >= 10) %>%
  count(performer) %>%
  filter(n >= 30) %>%
  arrange(desc(n))

# Bar plot for performers with at least 30 ten-week hits
ggplot(ten_week, aes(x = reorder(performer, n), y = n)) +
  geom_bar(stat = "identity", fill = "aquamarine") +
  labs(title = "Artists with at Least 30 'Ten-Week Hits'",
       x = "Performer",
       y = "Number of Ten-Week Hits") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The number of 'Ten-Week Hits' for artists having at least 30 songs that have been ten weeks or longer on the Billboard Top 100 chart is displayed in this bar plot. Every bar signifies a performer, and the height of the bar indicates how many songs satisfy the requirements. The artists are arranged according to the quantity of "Ten-Week Hits" in descending order. The artist with the most is Elton John, going above 50, while artists like Brad Paisley and those above him hover in the 30-40 range.